An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2023-09-19 arXiv/article.tex
added submission to arXiv

\documentclass[article, shortnames]{jss}
%% -- LaTeX packages and custom commands ---------------------------------------

\usepackage[usenames,dvipsnames]{xcolor}

@@

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
    \end{center}
    \caption[Baseline example for creating an \pkg{ELFI} model]{Baseline example for creating an \pkg{ELFI} model. Image taken from \cite{1708.00707}}
    \label{fig:elfi}
\end{figure}

@@

\input{latex_files/graphs/romc_graph}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_gt.pdf}
    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_posterior.pdf}
    % \input{latex_files/images/chapter3/example_gt.tex}
    % \input{latex_files/images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualization of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.pdf}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.pdf}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimization problem. In the left figure the region obtained with gradient-based optimizer and in the right one with Bayesian Optimization.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.pdf}
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimizer
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      \(\mu\) and the horizontal (black) the standard deviation
      \(\sigma\).}
  \label{fig:ma2_3}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.pdf}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.pdf}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_nn.pdf}
    \end{center}
    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
  \label{fig:ma2_4}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_solve.pdf}
    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_regions.pdf}
    \end{center}
    \caption[Execution time exploiting Parallelize]{Comparison
      between parallel and sequential execution of ROMC. We observe
      that the parallel version runs almost 5 times faster.}
      \label{fig:exec_parallel}
\end{figure}

@@

The results in this paper were obtained using \proglang{Python~3.9},
\pkg{ELFI~0.8.4} at \proglang{Ubuntu 20.04 lts} operating system, at a
single machine with \proglang{Intel® Core™ i7-8750H} processor.

>>>>

\documentclass[nojss]{jss}
%% -- LaTeX packages and custom commands ---------------------------------------
\pdfoutput=1
\usepackage[usenames,dvipsnames]{xcolor}

@@

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.8\textwidth]{./images/chapter2/elfi.png}
    \end{center}
    \caption[Baseline example for creating an \pkg{ELFI} model]{Baseline example for creating an \pkg{ELFI} model. Image taken from \cite{1708.00707}}
    \label{fig:elfi}
\end{figure}

@@

\input{graphs/romc_graph}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{images/chapter3/example_gt.pdf}
    \includegraphics[width=0.49\textwidth]{images/chapter3/example_posterior.pdf}
    % \input{images/chapter3/example_gt.tex}
    % \input{images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualization of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1.pdf}
        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1_bo.pdf}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimization problem. In the left figure the region obtained with gradient-based optimizer and in the right one with Bayesian Optimization.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_nn.pdf}
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimizer
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      \(\mu\) and the horizontal (black) the standard deviation
      \(\sigma\).}
  \label{fig:ma2_3}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior.pdf}
    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_bo.pdf}
    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_nn.pdf}
    \end{center}
    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
  \label{fig:ma2_4}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_solve.pdf}
    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_regions.pdf}
    \end{center}
    \caption[Execution time exploiting Parallelize]{Comparison
      between parallel and sequential execution of ROMC. We observe
      that the parallel version runs almost 5 times faster.}
      \label{fig:exec_parallel}
\end{figure}

@@

The results in this paper were obtained using \proglang{Python~3.9},
\pkg{ELFI~0.8.6} at \proglang{Ubuntu 20.04 lts} operating system, at a
single machine with \proglang{Intel® Core™ i7-8750H} processor.

### 2023-09-19 arXiv/graphs/romc_graph.tex
added submission to arXiv

      \end{tikzpicture}
    }
  \end{center}
  \caption{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses \textbf{There are no ellipses.} must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
\label{fig:romc_overview}
\end{figure}

>>>>

      \end{tikzpicture}
    }
  \end{center}
  \caption{Overview of the ROMC implementation. On the left side, we
    depict ROMC as a sequence of algotirhmic steps. On the right side,
    we present the functions that form our implementation; the green
    rectangles (starting with underscore) are the internal
    functionalities and the blue rectangles the publicly exposed
    API. This side-by-side illustration highlights that our
    implementation follows strictly the algorithmic view of ROMC.}
\label{fig:romc_overview}
\end{figure}
